\documentclass{article}

\title{Travel Grants Applying Position Paper}
\author{Mingliang Liu}
\date{26 June, 2011}

\begin{document}
\maketitle

I would like to attend the APSys'11 for two reasons. First, I shall present
our paper in the oral session. The paper is named \textit{One Optimized I/O Con-
figuration per HPC Application: Leveraging the Configurability of Cloud }. Our
exploration indicates that customized system configuration for HPC
applications in the cloud is important and non-trivial. Second, I'm looking
forward to listen the latest works the other groups done, especially in the
Asian area. I read most of the papers published last year at the first APSys
proceeding, and found them interesting.

I focus myself on the parallel I/O systems in the cloud platforms such Amazon
EC2. The storage system options in the cloud behave differently from the local
clusters in many ways. I did some preliminary work and would like to show them
in the conference. Moreover, discussing with other guys who attend the meeting
should give me some useful hints too. In the future, I would like to
automating the I/O system configuration process in the cloud for individual
HPC applications. I believe the attendance of the APSys'11 would help me to
gain more insights into the HPC in cloud topic.

\vspace{.5cm}
Due to the time limit, I shall take the airplane to join the meeting from
Beijing. It will cost me about \$250 according to the latest discount
information. The accommodation costs is about \$100 for three days. The daily
expense covering meal costs and taxi is predicted \$50. There is a potential
cost to run sample experiments during the conference in the Amazon EC2 cloud.
This may happen if someone is interested in our customized Amazon Machine
Instances, file system building scripts and so on. I think the range of this
cost is between \$50 and \$100. Totally, I apply \textbf{\$450} for my trip.

\vspace{.5cm}
Thank you for consideration.

\end{document}
